Finite Property Convex Elicitation Paper

### 2019-01-26 cvx-fin.tex
tidy up 1d

\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\begin{proposition}\label{prop:optimal-reports-per-level-set}
  \raft{Jessie: I think this should be fine for the paper.  We may want strip and trim for the proof, in the appendix, in which case we can define them there...}
  Let $\Gamma:\simplex\toto\reals^d$ be a non-degenerate (convex) elicitable property.
  \raf{Do we really need $\Gamma$ to be convex elicitable?}
  \jessie{Not a requirement... I threw it in parentheses since we typically think of $\Gamma$ as elicited by a convex surrogate, but it isn't used in the proof.}

  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite elicitable property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R)\} =: \trim(\Gamma)|_{\varphi(\R)}$.
  Clearly, $\trim(\Gamma) \supseteq \trim(\Gamma)|_{\varphi(\R)}$ as we are simply restricting the number of level sets we are indexing over.
  Now, consider some $\theta \in \trim(\Gamma)$ and the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly we are done.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done since the level sets composing $\trim$ are unlabeled.
  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  If $\Gamma_u$ is full-dimensional, it must be unique up to relabeling, giving us one of the first two cases or a contradiction of elicitability.
  If $\Gamma_u$ is not full-dimensional, it must be a subset of some full-dimensional level set, and therefore not in $\trim(\Gamma)$.
  Thus, we conclude $\trim(\Gamma) \subseteq \trim(\Gamma)|_{\varphi(\R)}$ and since the sets are equal, can conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

@@

\begin{proposition}
  If $\Gamma$ embeds $\gamma$, then there is a link function $\psi$ between $\Gamma$ and $\gamma$.
  Moreover, if $L$ elicits $\Gamma$ and $L$ is polyhedral, $\psi$ can be taken to be separated.
\end{proposition}
\begin{proof}

%Let $\varphi$ be the embedding function.
%Then the function $\psi(u) := \varphi^{-1}(r)$ is a link between $\Gamma$ and $\gamma$.
%Since, for all $p \in\simplex$, we know $\varphi(r) \in \Gamma(p) \implies \varphi^{-1}(\varphi(r)) = r \in \gamma(p)$, we can observe that $\psi$ is a calibrated link.
%
%For the second part, in order for $\psi$ to be separated, for each embedded point $u \in \reals^d$ and every $v \in B(\epsilon, u)$, we must observe $\psi(v) = \psi(u)$.
%If the loss $L$ is polyhedral, then we can construct a $\psi$ so that the above holds.
%If no such $\psi$ existed, then for some $p$ such that $u \in \arginf_{z \in \R} \E_p L(z,Y)$, we know there is a $\delta > 0$ such that $\E_p L(v,Y) - \E_p L(u,Y) < \delta$ by continuity of $L$.
%Additionally, if $\psi$ was not selective, then we know $\psi(v) = \psi(w)$ for some embedded point $w \neq u$.
%Since the losses of $u$ and $v$ are not bounded about away from each other and $\psi(v) = \psi(w)$, then the loss $L$ must be constant on $\conv(\{u,v,w\})$, otherwise we would either contradict elicitability of $\Gamma$ or the construction of the calibrated link $\psi$.
%Therefore, we can reassign $\psi(v) = \psi(u)$ for any $v\in B(\epsilon, u)$, and such a link additionally separated.
\end{proof}

@@

\begin{theorem}
  Every polyhedral loss embeds a finite elicitable property.

  $+$

  A polyhedral loss indirectly elicits a finite elicitable property $\gamma$ if and only if it embeds a property which refines $\gamma$.
\end{theorem}

\raf{But wait, what about dimension?}

@@

\begin{definition}
  The \emph{embedding dimension} of a finite property $\gamma$, written $\elicembed(\gamma)$, is the minimum $d\leq 0$ \jessie{$d \geq 1?$} such that $\gamma$ is $d$-embeddable.
\end{definition}

@@

\raf{Note: move monotone and lemmas to appendix}

@@

\begin{proposition}\label{prop:indirect-orderable}
  If convex $L : \reals \to \reals^\Y$ indirectly elicits a finite elicitable property $\gamma$, then $\gamma$ is orderable.
\end{proposition}
\begin{proof}
  Let $\gamma:\simplex\toto\R$.
  From Lemma~\ref{lem:prop-L-monotone}, $\prop{L}$ is monotone.
  Let $\psi:\reals\to\R$ be the calibrated link from $\prop{L}$ to $\gamma$.

\raf{Now follows from the previous and polyhedral $\implies$ embeds}
\begin{corollary}\label{cor:embed-orderable}
  Every polyhedral $L : \reals \to \reals^\Y$ embeds an orderable property.
\end{corollary}

\begin{proposition}
  Every orderable property $\gamma$ is 1-embeddable.
\end{proposition}
\begin{proof}
  Let $\R = \{r_1,\ldots,r_k\}$ be the report space for $\gamma$.
  From Lemma~\ref{lem:prop-L-monotone}, and its proof, we have $a:\R\to\reals^\Y, b\to\reals^\Y$ such that $\gamma_r = \{p\in\simplex : a(r) \cdot p \leq 0 \leq b(r) \cdot p\}$ for all $r\in\R$, and for all $i \in \{1,\ldots,k-1\}$, we have $a(r_i) \leq b(r_i) = a(r_{i+1}) \leq b(r_{i+1})$.
  We now define $\varphi(r_i) = i \in \reals$, and $L$... \raf{integrate?  this should be simple... want $\partial L(i)_y = [a(r_i)_y,b(r_i)_y]$.}
\end{proof}

We can summarize the above in the following theorem.
\raf{TFAE}
\begin{theorem}
  Let $\gamma$ be a finite elicitable property.
  Then $\gamma$ is 1-embeddable if and only if it is orderable.
  In particular, if $\eliccvx(\gamma)=1$, then $\elicpoly(\gamma) = \elicembed(\gamma)=1$.
  \raf{add $elic$ for continuous non-loc-const properties}
\end{theorem}

@@

>>>>

\newcommand{\prop}[1]{\Gamma[#1]}
\newcommand{\eliccts}{\mathrm{elic}_\mathrm{cts}}
\newcommand{\eliccvx}{\mathrm{elic}_\mathrm{cvx}}
\newcommand{\elicpoly}{\mathrm{elic}_\mathrm{pcvx}}
\newcommand{\elicembed}{\mathrm{elic}_\mathrm{embed}}

@@

\begin{proposition}\label{prop:embed-trim}
  Let $\Gamma:\simplex\toto\reals^d$ be an elicitable property.
  The following are equivalent:
  \begin{enumerate}
  \item $\Gamma$ embeds a finite elicitable property $\gamma:\simplex \toto \R$.
  \item $\trim(\Gamma)$ is a finite set.
  \item There is a finite set of full-dimensional level sets $\Theta$ of $\Gamma$ that union to $\simplex$.
  \end{enumerate}
  Moreover, when any of the above hold, $\{\gamma_r : r\in\R\} = \trim(\Gamma) = \Theta$.
\end{proposition}
\begin{proof}
  For $1 \implies 2$, since $\Gamma$ embeds the finite property $\gamma$, we show that $\trim(\Gamma) = \{\Gamma_u : u \in \varphi(\R)\} =: \trim(\Gamma)|_{\varphi(\R)}$.
  Clearly, $\trim(\Gamma) \supseteq \trim(\Gamma)|_{\varphi(\R)}$ as we are simply restricting the number of level sets we are indexing over.
  Now, consider some $\theta \in \trim(\Gamma)$ and the report $u$ so that $\theta = \Gamma_u$.
  If $u \in \varphi(\R)$, then clearly we are done.
  If $u \not\in \varphi(\R)$ but there is some report $w \in \varphi(\R)$ so that $\Gamma_u = \Gamma_w$, then we are again done since the level sets composing $\trim$ are unlabeled.
  If there is no such $w \in \R$, then we claim $\Gamma_u \not \in \trim(\Gamma)$.
  If $\Gamma_u$ is full-dimensional, it must be unique up to relabeling, giving us one of the first two cases or a contradiction of elicitability.
  If $\Gamma_u$ is not full-dimensional, it must be a subset of some full-dimensional level set, and therefore not in $\trim(\Gamma)$.
  Thus, we conclude $\trim(\Gamma) \subseteq \trim(\Gamma)|_{\varphi(\R)}$ and since the sets are equal, can conclude that if $\Gamma$ embeds the finite property $\gamma$, then $\trim(\Gamma)$ is finite.

@@

\begin{proposition}\label{prop:embed-link}
  If $\Gamma$ embeds a finite property $\gamma$, then it has a calibrated link $\psi$ to $\gamma$.
  Moreover, if a polyhedral loss $L$ elicits $\Gamma$, $\psi$ can be taken to be separated with respect to $L$.
\end{proposition}
\begin{proof}
Let $\varphi$ be the embedding function.
Then the function $\psi(u) := \varphi^{-1}(r)$ is a link between $\Gamma$ and $\gamma$.
\raf{I think you mean $\varphi^{-1}(u)$, but that is only defined on $\varphi(\R)$ (the embedded points).  Use Proposition~\ref{prop:embed-trim}, specifically trim.}
Since, for all $p \in\simplex$, we know $\varphi(r) \in \Gamma(p) \implies \varphi^{-1}(\varphi(r)) = r \in \gamma(p)$, we can observe that $\psi$ is a calibrated link.

For the second part, in order for $\psi$ to be separated, for each embedded point $u \in \reals^d$ and every $v \in B(\epsilon, u)$, we must observe $\psi(v) = \psi(u)$.
If the loss $L$ is polyhedral, then we can construct a $\psi$ so that the above holds.
If no such $\psi$ existed, then for some $p$ such that $u \in \arginf_{z \in \R} \E_p L(z,Y)$, we know there is a $\delta > 0$ such that $\E_p L(v,Y) - \E_p L(u,Y) < \delta$ by continuity of $L$.
Additionally, if $\psi$ was not selective, then we know $\psi(v) = \psi(w)$ for some embedded point $w \neq u$.
Since the losses of $u$ and $v$ are not bounded about away from each other and $\psi(v) = \psi(w)$, then the loss $L$ must be constant on $\conv(\{u,v,w\})$, otherwise we would either contradict elicitability of $\Gamma$ or the construction of the calibrated link $\psi$.
Therefore, we can reassign $\psi(v) = \psi(u)$ for any $v\in B(\epsilon, u)$, and such a link additionally separated.
\end{proof}

@@

\begin{theorem}\label{thm:polyhedral-embed}
  Every polyhedral loss embeds a finite elicitable property.
  Moreover, a polyhedral loss indirectly elicits a finite elicitable property $\gamma$ if and only if it embeds a property which refines $\gamma$.
\end{theorem}

\raf{Transition: wait, what about dimension?}

@@

\begin{definition}
  The \emph{embedding dimension} of a finite property $\gamma$, written $\elicembed(\gamma)$, is the minimum $d\geq 0$ \jessiet{$d \geq 1?$}\raft{Technically constant properties $\gamma(p) = \{c\} \forall p$ would have complexity 0.  Not terribly interesting... but it would also feel weird to call that complexity $1$.  Hmm.} such that $\gamma$ is $d$-embeddable.
\end{definition}

@@

@@

Lambert et al. shows that a finite property has a \emph{order-sensitive} loss, meaning one assigning higher loss to ``farther'' reports (with respect to a given order), if and only if it is orderable.
We will see that, in fact, a finite property is 1-embeddable if and only if it is orderable.
First, the forward direction.

\begin{proposition}\label{prop:orderable-embed}
  Every orderable property $\gamma$ is 1-embeddable.
\end{proposition}
\begin{proof}
  Let $\R = \{r_1,\ldots,r_k\}$ be the report space for $\gamma$.
  From Lemma~\ref{lem:prop-L-monotone}, and its proof, we have $a:\R\to\reals^\Y, b\to\reals^\Y$ such that $\gamma_r = \{p\in\simplex : a(r) \cdot p \leq 0 \leq b(r) \cdot p\}$ for all $r\in\R$, and for all $i \in \{1,\ldots,k-1\}$, we have $a(r_i) \leq b(r_i) = a(r_{i+1}) \leq b(r_{i+1})$.
  We now define $\varphi(r_i) = i \in \reals$, and $L$... \raf{integrate?  this should be simple... want $\partial L(i)_y = [a(r_i)_y,b(r_i)_y]$.}
\end{proof}

The converse of Proposition~\ref{prop:orderable-embed} is a corollary of the following result, together with Theorem~\ref{thm:polyhedral-embed} (polyhedral losses embed a finite elicitable property) and Proposition~\ref{prop:embed-link} (embeddings give rise to calibrated links).
For a proof, see Appendix~\ref{app:dimension-1}.
\begin{proposition}\label{prop:indirect-orderable}
  If convex $L : \reals \to \reals^\Y$ indirectly elicits a finite elicitable property $\gamma$, then $\gamma$ is orderable.
\end{proposition}

\begin{corollary}\label{cor:embed-orderable}
  Every polyhedral $L : \reals \to \reals^\Y$ embeds an orderable property.
\end{corollary}

Combining these above observations with a few others, we arrive at the following result, proven in Appendix~\ref{app:dimension-1}, which states that several elicitation complexity classes collapse in dimension 1.
\begin{theorem}
  Let $\gamma$ be a finite elicitable property.
  The following are equivalent.
  \begin{enumerate}
  \item $\gamma$ is orderable.
  \item $\elicembed(\gamma)=1$. ($\gamma$ is 1-embeddable.)
  \item $\elicpoly(\gamma)=1$. ($\gamma$ is indirectly elicitable via a polyhedral loss $L:\reals\to\reals^\Y$.)
  \item $\eliccvx(\gamma)=1$. ($\gamma$ is indirectly elicitable via a convex loss $L:\reals\to\reals^\Y$.)
  \item $\eliccts(\gamma)=1$. ($\gamma$ is indirectly elicitable via a real-valued continuous non-locally-constant property.)
  \end{enumerate}
\end{theorem}

@@

\appendix

\section{Dimension 1}\label{app:dimension-1}

We now prove Proposition~\ref{prop:indirect-orderable}, which states the following: if convex $L : \reals \to \reals^\Y$ indirectly elicits a finite elicitable property $\gamma$, then $\gamma$ is orderable.
\begin{proof}[of Proposition~\ref{prop:indirect-orderable}]
  Let $\gamma:\simplex\toto\R$.
  From Lemma~\ref{lem:prop-L-monotone}, $\prop{L}$ is monotone.
  Let $\psi:\reals\to\R$ be the calibrated link from $\prop{L}$ to $\gamma$.
